---
title: "PassSanitaire"
author: "FD"
output:
  bookdown::html_document2:
    self_contained: no
    keep_md: true
    number_sections: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

load("../code/savevalues.RData")
load("../code/data_indicators.RData") 
```

The French domestic Pass Sanitaire did not solve vaccination inequities: a nationwide longitudinal study on `r round(totPopSize / 10^6)` million individuals


# Abstract

Context: Implemented during the Summer 2021, the French "Sanitary Pass" led to an increase in vaccination rates in France, but spatial heterogeneities in vaccination rates are still striking [LG1]. It morphed into a "Vaccine Pass" in early 2022. To identify potential determinants of these heterogeneities and evaluate the French Sanitary and Vaccine Pass' efficacies in reducing them, we used a data-driven approach on exhaustive nationwide data, gathering `r nIndicatorsAfterRemovingCorrMiss` socio-economic, political and geographic indicators.

Methods: We considered the association between being a district above the median value of the first-dose vaccination rates and being above the median value of each indicator at different time points: just before the sanitary pass announcement (week `r format(as.Date(datesAnalysis[1]), "%Y-W%W")`), just before the sanitary pass came into force (week `r format(as.Date(datesAnalysis[2]), "%Y-W%W")`) and one month after (week `r format(as.Date(datesAnalysis[3]), "%Y-W%W")`), and the equivalent dates for the vaccine pass (weeks `r format(as.Date(datesAnalysis[4:6]), "%Y-W%W")`). We then considered the change over time of vaccination rates according to the XX most associated indicators, by deciles. 

Results: The XX indicators most associated with vaccination rates were XX, XX, XX. 

Conclusion: Our analysis reveals that, both before and after the introduction of the French sanitary pass, factors with the largest impact are related to poverty XX, XXX, XXX. 

# Introduction

The rapid development of effective COVID-19 vaccines brought the hope of a rapid return to normalcy, but heterogeneous vaccination rates, both among countries because of inequitable distributions of doses [Usher21] and within countries [Caspi21, Murthy21], jeopardize epidemic control.
 
Hesitancy and hostility toward vaccination have been comparatively high in France in recent decades. It is particularly the case since the 2009-2010 vaccination campaign against pandemic flu, whose contested management in France was a tipping point that led to higher vaccine hesitancy and hostility [Ward19; Guimier2021]. 
The trend was confirmed with the COVID-19 pandemic [Spire21]: just before Covid-19 vaccines became available, intentions to get vaccinated were comparatively very low in France (44% of the respondents in [Wouters2021] in the Fall 2020; about 40% of respondent in [SantePubliqueFrance21] in December 2020). Acceptance of the COVID-19 vaccine however gradually grew during 2021 [SantePubliqueFrance22].
<!-- Lazarus21 does not talk directly about France --> 
 
Spatial heterogeneties in vaccination rates have already been documented in France for previous vaccines. Vaccination coverage for the Hepatitis B vaccine and for the Measles-Mumps-Rubella vaccine has been lower in the South of France, and especially in the South-East of the country [Guimier21]. Distance to the central political power in Paris, as well as a sense of belonging to a local community with a strong cultural identity, have been put forward as potential explanations for this geographic gradient in vaccination rates [Guimier21]. 

Attitudes toward vaccination are also known to be influenced by social and territorial inequalities. Surveys conducted in 2020 in France showed that respondents with lower education [Schwarzinger21, Spire21], lower income levels or less trust in authorities [Spire21] were more likely to be hostile to COVID-19 vaccines. A survey conducted in July 2021 in France [Bajos2022] on close to 81000 participants indicated that unvaccinated respondents were more likely to have lower income and more likely to belong to racialised minorities than vaccinated ones. The survey also confirmed that mistrust in the government was fueling refusal to be vaccinated [Bajos2022]. 

<!-- XX MOVE INTERNATIONAL COMPARISONS IN THE DISCUSSION Whether differential intentions translated into effective differences in vaccination, as described in other countries [Caspi21, Murthy21], remained to be shown in France. (XX New Bajos Paper?) -->
 
By mid-July 2021, XX% of the eligible population in France had received a first dose, and the France was facing an epidemic wave due to the Delta variant. To speed up vaccination, President Macron announced on 12 July 2021 the implementation of a domestic "sanitary pass" (le passe sanitaire), which came fully into force on 9 August 2021. The "sanitary pass" was required in most cultural venues, for both indoor and outdoor dining and in health structures. This announcement led to an unprecedented demand for vaccination, which was celebrated internationally [Ref NYT?]. Vaccination rates climbed from 54% of the whole population (XX change into eligible) by 12 July 2021 to 69% on 4 September 2021. Because it targeted pay-for social activities, however, the "sanitary pass" was feared to have a limited impact on vaccination inequities.

This study aims to obtain further insights on the socio-economic, political and geographic factor associated with vaccination rates, and to evaluate the effect of the French domestic sanitary pass, by using nation-wide, exhaustive datasets.
 
# Methods

## Data 

<!-- Vaccination -->
The French state health insurance service (Assurance Maladie) provides public datasets of vaccination rates in France. These datasets are based on aggregated individual data on beneficiaries of the national health insurance service who received health care in the past year. These exhaustive datasets are updated weekly, and are provided at the district scale nationally (EPCI: *Établissement public de coopération intercommunale*, an administrative level gathering multiple towns or cities) and at the suburban scale for the Paris, Lyon, and Marseille metropolitan areas. For this study, we focused on Metropolitain France, because vaccination rates are much lower in oversea localities, and because determinants of vaccination rates are likely to differ in oversea localities compared to metropolitain ones. Our dataset included `r sum(nlocalities)` districts (`r nlocalities["EPCI"]` EPCI and `r nlocalities["PLM"]` districts at the suburban scale in Paris, Lyon, Marseille). 

The vaccination dataset for mainland France encompasses about XX individuals (median district size XX, itq range XX). The vaccination data are available by age class: 00--19, 20--39, 40--54, 55--64, 65--74, 75 and over. Population sizes for each locality and each age class are also provided. These population sizes are updated monthly to account for individuals aging, potential residential moves, and deaths. 


<!-- Other data -->

```{r, echo = FALSE}
insee.indicators.names <- names(typesDataAll)[!is.element(names(typesDataAll), c("Geography", "Elections"))]

tb <- typesDataNoCorr[!is.element(names(typesDataNoCorr), c("Geography", "Elections"))]
output.INSEE <- ""
for(i in 1:length(tb)) output.INSEE <- paste0(output.INSEE, names(tb)[i], ": n = ", tb[i], "; ")
output.INSEE <- substr(output.INSEE, 1, nchar(output.INSEE)-2)
```

We paired these vaccination data with three other datasets gathering socio-economic, political and geographic variables.

Socio-economic data are provided by the French national statistics institute (INSEE), and are available at the same administrative levels as the vaccination data. The different predictors are classified into `r length(insee.indicators.names)` categories (`r insee.indicators.names`).  
Latitude, longitude and surface data were extracted from maps datasets [ref OpenMaps]. We calculated from them four additional geographic indicators: distance to Paris, relative position along a South-East--North-West gradient, relative position along a South-West--North-East gradient, and local population density.  
Our political dataset consisted of the results of the two round of the 2017 Presidential election in France, which we aggregated to reconstitute the same administrative levels as the vaccination dataset. This political dataset contains the proportions of votes for each of the 11 candidates of the first round, 2 candidates of the second round (Macron and Le Pen), and proportion of abstention at each round. 

These three datasets comprised `r nIndicatorsTot` indicators. We then removed those indicators with over `r round(thrMissingVals * 100)`% missing data, or with over `r thrCor` correlation with other indicators of the dataset, which left us with `r nIndicatorsAfterRemovingCorrMiss` indicators: `r sum(tb)` socio-economic indicators (`r output.INSEE`); `r typesDataNoCorr[is.element(names(typesDataNoCorr), c("Geography"))]` geographic indicators; `r typesDataNoCorr[is.element(names(typesDataNoCorr), c("Elections"))]` political indicators. 

## Analysis

Vaccination was accessible to all adults in France after 27 May 2021. It opened to teenagers (12-17 year olds) on 15 June 2021, and to children (5-11 year olds) on 22 December 2021. Because of this differential accessibility of vaccines, and because vaccine passport rules also differed for non-adults, we excluded the 00-19 age class from our analysis, and focused on vaccination rates among 20+ year-old individuals (hereafter "adults"). 

For each indicator in our dataset, at each of the four chosen dates (weeks `r format(as.Date(datesAnalysis), "%Y-W%W")`), we considered the association between living in a district above the median of a that indicator and individual first-dose vaccination rates among adults. Odds ratios (OR) were computed from the output of a logistic regression. To be able to compare predictors irrespective of the direction of the effect, we considered the maximum of `OR, 1/OR` (hereafter $\overline{OR}$). Vaccination data are at the individual level, and indicator data at the district level.  
For each date, we determined a significance threshold by computing odds ratios on `r nrep` random permutations of a predictor, and identifying the value of the XX 99% percentile odd ratios ($\overline{OR}$) of these permuted data. 

For representative indicators with the strongest association to vaccination data, we estimated age-adjusted vaccination rates among adults over time, for each decile of each indicator (treated as a factor). These estimations were obtained from a logistic model taking age class into account, and adult vaccination rates were computed on a standardized age distribution matching that of metropolitain France. 

All analysis code is available at XXX; analyses were done in `r R.Version()$version.string`. 



<!-- 

- **27 mai 2021 (S21)** : toutes les personnes majeures peuvent prendre rendez-vous pour se faire vacciner en centre de vaccination pour des injections avec les vaccins Pfizer-BioNTech ou Moderna à compter du 31 mai 2021.
- **15 juin 2021 (S24)** : les 12 à 17 ans inclus ans peuvent se faire vacciner avec le vaccin Pfizer-BioNTech en centre de vaccination. Le 24 juillet 2021, l'Agence européenne du médicament (EMA) a délivré une autorisation du vaccin Moderna sur lequel la HAS a rendu un avis favorable le 28 juillet.
- **12 juillet 2021 (S28)** : annonce du président le la république concernant la mise en place du pass sanitaire.
- **9 août 2021 (S32)** : mise en place de l'obligation du passe sanitaire dans les lieux concernés.
- **1er septembre 2021 (S35)** : une campagne de rappel est mise en place pour stimuler le système immunitaire des plus vulnérables : les résidents d'EHPAD et d'USLD, les personnes de 65 ans et plus vivant à domicile, les personnes souffrant de comorbidité(s), les personnes atteintes de pathologies à haut risque de forme grave, les personnes ayant été vaccinées avec le vaccin Janssen.
- **15 septembre 2021 (S37)** : l'obligation vaccinale concernerait tous les personnels des établissements de santé, établissements d'hébergement pour personnes âgées dépendantes (Ehpad) et établissements pour personnes en situation de handicap et tous les professionnels ou les bénévoles en contact avec des personnes âgées ou vulnérables, y compris à domicile. Cette mesure a fait l'objet d'un projet de loi relatif à la gestion de la crise sanitaire examiné par le Parlement depuis le 20 juillet 2021.


-->


# Results

We investigated the associations between each of the `r nIndicatorsAfterRemovingCorrMiss` indicators and first-dose vaccination rates. Two indicators were among the top five most associated one at all time points (see Figure \@ref(fig:figManhattan)): the share of local income coming from unemployment benefits (`Unemployment_Benef`) and vote for Mr Asselineau (`Asselineau`). The three other most associated indicators did not change in the later dates that we considered, and were the proportion of immigrants in the district (`Immigrant`), the district's relative position along a North-West--South-East gradient (`NO-SE`) and the proportion of overcrowded household (`Overcrowding_rate`). 

GIVE NAMES OF NOT ASSOCIATED INDICATORS?

GIVE VALUES IN A TABLE?

```{r figManhattan, echo=FALSE, fig.height=9, fig.width=9, fig.cap="Manhattan plots of the Odds ratios for each of the indicator of our dataset, by date. Left column: around the Sanitary Pass implementation; right column: around the Vaccine Pass implementation. The top five odds ratios are labelled at each time point; the arrow next to the name indicates the direction of the effect. The gray rectangle corresponds to the 99% percentile of odds ratios in the permuted data; points falling in the rectangle are considered as non-significant."}
load("../data/predictors.RData")
load("../code/outLR.RData")
source("../code/2_plot-manhattan.R")
layout(matrix(1:6, ncol = 2))
# PLOT
plotManhattan(outLR, ntop = 5)
```

Our odds ratio calculations were based on a crude version of each indicator, which were dichotomized into values above or below the median of each indicator. To better visualize the effects (or lack thereof) of the sanitary and vaccine passes on vaccination rates over time, we computed age-adjusted vaccination rates over time, by decile of three of the most associated indicators, treated as factors (see Figure \@ref(fig:figOverTime)). While the Sanitary Pass led to an overall increase in vaccination rates, the Vaccine Pass did not affect the trend of change in vaccination rates. Heterogeneities in vaccination rates persisted after both types of pass, neatly ordered by decile of each indicator. 


```{r figOverTime, fig.height=10, fig.width=4, fig.cap="Age-adjusted vaccination rates among adults, over time, by decile of each indicator. ", echo = FALSE}
load("../code/outTime.RData")
source("../code/2_plot-overTime.R")
layout(matrix(1:3, ncol = 1))
plotPropTime(outA, plotDates = TRUE, plotGraduations = TRUE)
```

Finally, historically under-vaccinated areas in France stand out as being less vaccinated against Covid-19, in particular the South-East region.  

FIGURE MAP


# Discussion

